---
title: "analysis-planning"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(ggplot2)
require(dplyr)
```

## Study summary

- **N**=33 participants
- Sorted 20 exemplars of grayscale wallpapers
- P1, P3M1, P31M, P6M, P6M

## Questions

1. Are some wallpaper groups more similar to one another?
    - Fewer sets
    - Higher 'similarity' ratings among exemplar pairs
2. Are some exemplars more representative than others?

## Exemplars per set?

<div class="centered">
```{r read-plot-nsets, echo=FALSE, message=FALSE, warning=FALSE}
n.sets.df <- read.csv('analysis/data/nsets.csv', header=TRUE)

theme.symm <- theme( axis.title.x = element_text(size=18),
                       axis.title.y = element_text(size=18),
                       strip.text = element_text(size=16),
                       axis.text = element_text(size=14),
                       legend.text = element_text(size=14),
                       legend.title = element_text(size=14) )

p <- ggplot(data = n.sets.df)
p + 
  aes(x=Group, y = Nsets, color=Group) + 
  geom_violin() + geom_jitter(width = .5, height=0, alpha=.3)
```
</div>

## Individual differences exemplars per set

<div class="centered">
```{r n.sets-by-participant}
qplot(x=Participant, y=Nsets,
      data=n.sets.df,
      geom=c("point"),
      facets = Group ~ ., fill=Group, color=Group ) +
  labs( x="Participant", y="N observations") +
  theme( axis.text.x = element_blank() )+
  guides( fill=FALSE )
```
</div>

## Jaccard index as a metric of similarity/dissimilarity

- <https://en.wikipedia.org/wiki/Jaccard_index>
- $J(A,B)=\frac{A \cap B}{A \cup B}$
- $0 <= J(A,B) <= 1$

## Jaccard indices by wallpaper group

<div class="centered">
```{r read-plot-jaccard-by-group, echo=FALSE, message=FALSE, warning=FALSE}
jaccard.df <- read.csv(file='analysis/data/jaccard.csv', header=TRUE)

qplot(x=Jaccard,
      data=jaccard.df,
      geom=c("histogram"),
      facets = Group ~ ., fill=Group ) +
  labs( x="Jaccard Index", y="N observations") +
  theme.symm +
  guides( fill=FALSE )
```
</div>

## Jaccard indices by exemplar

<div class="centered">
```{r read-plot-jaccard-by-exemplar, echo=FALSE, message=FALSE, warning=FALSE}
# Re-read data files to includes a,b and b,a scores for each
# exemplar pair. Of course j(a,b)=j(b,a), but we need both scores to show
# exemplar-by-exemplar patterns. In the future, just do this then filter.

jaccard.all.df <- read.csv(file='analysis/data/jaccard.csv', header=TRUE)

# Select a,b and b,a but not a,a pairs
jaccard.select <- with(jaccard.all.df, !(Exemplar.Row==Exemplar.Col))

qplot(data=jaccard.all.df[jaccard.select,],
      x=as.factor(Exemplar.Row), y=Jaccard,
      facets = Group ~ .,
      fill=Group,
      geom="boxplot",
      xlab = "Exemplar")
```
</div>

## Exemplar pairs with high Jaccard indices

```{r top-Jaccard-indices, echo=FALSE, message=FALSE, warning=FALSE}
max.jaccard.df <- jaccard.df %>% 
  group_by(Group) %>% 
  summarise(max.jaccard = max(Jaccard))

max.jaccard.merged.df <- merge(max.jaccard.df, jaccard.df, by="Group")

max.jaccard.merged.df <- max.jaccard.merged.df %>%
  filter(Jaccard == max.jaccard) %>%
  select(Exemplar.Row, Exemplar.Col, Jaccard, Group)

next.group.i <- 1
```

## J(`r max.jaccard.merged.df$Exemplar.Row[next.group.i]`, `r max.jaccard.merged.df$Exemplar.Col[next.group.i]`; `r max.jaccard.merged.df$Group[next.group.i]`)=`r 
max.jaccard.merged.df$Jaccard[next.group.i]`

<div class="centered">
```{r show-1-pair, echo=FALSE, warning=FALSE, message=FALSE}
source('analysis/display.exemplar.pairs.R')
display.exemplar.pairs(max.jaccard.merged.df$Exemplar.Row[next.group.i], max.jaccard.merged.df$Exemplar.Col[next.group.i], max.jaccard.merged.df$Group[next.group.i])
next.group.i <- 2
```
</div>

## J(`r max.jaccard.merged.df$Exemplar.Row[next.group.i]`, `r max.jaccard.merged.df$Exemplar.Col[next.group.i]`; `r max.jaccard.merged.df$Group[next.group.i]`)=`r 
max.jaccard.merged.df$Jaccard[next.group.i]`

<div class="centered">
```{r show-2-pair, echo=FALSE, warning=FALSE, message=FALSE}
source('analysis/display.exemplar.pairs.R')
display.exemplar.pairs(max.jaccard.merged.df$Exemplar.Row[next.group.i], max.jaccard.merged.df$Exemplar.Col[next.group.i], max.jaccard.merged.df$Group[next.group.i])
next.group.i <- next.group.i + 1
```
</div>

## J(`r max.jaccard.merged.df$Exemplar.Row[next.group.i]`, `r max.jaccard.merged.df$Exemplar.Col[next.group.i]`; `r max.jaccard.merged.df$Group[next.group.i]`)=`r 
max.jaccard.merged.df$Jaccard[next.group.i]`

<div class="centered">
```{r show-3-pair, echo=FALSE, warning=FALSE, message=FALSE}
source('analysis/display.exemplar.pairs.R')
display.exemplar.pairs(max.jaccard.merged.df$Exemplar.Row[next.group.i], max.jaccard.merged.df$Exemplar.Col[next.group.i], max.jaccard.merged.df$Group[next.group.i])
next.group.i <- next.group.i + 1
```
</div>

## J(`r max.jaccard.merged.df$Exemplar.Row[next.group.i]`, `r max.jaccard.merged.df$Exemplar.Col[next.group.i]`; `r max.jaccard.merged.df$Group[next.group.i]`)=`r 
max.jaccard.merged.df$Jaccard[next.group.i]`

<div class="centered">
```{r show-4-pair, echo=FALSE, warning=FALSE, message=FALSE}
source('analysis/display.exemplar.pairs.R')
display.exemplar.pairs(max.jaccard.merged.df$Exemplar.Row[next.group.i], max.jaccard.merged.df$Exemplar.Col[next.group.i], max.jaccard.merged.df$Group[next.group.i])
next.group.i <- next.group.i + 1
```
</div>

## J(`r max.jaccard.merged.df$Exemplar.Row[next.group.i]`, `r max.jaccard.merged.df$Exemplar.Col[next.group.i]`; `r max.jaccard.merged.df$Group[next.group.i]`)=`r 
max.jaccard.merged.df$Jaccard[next.group.i]`

<div class="centered">
```{r show-5-pair, echo=FALSE, warning=FALSE, message=FALSE}
source('analysis/display.exemplar.pairs.R')
display.exemplar.pairs(max.jaccard.merged.df$Exemplar.Row[next.group.i], max.jaccard.merged.df$Exemplar.Col[next.group.i], max.jaccard.merged.df$Group[next.group.i])
next.group.i <- next.group.i + 1
```
</div>
