---
title: "jaccard-mds-analysis"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  ioslides_presentation:
    widescreen: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
require(ggplot2)
require(dplyr)
require(png)

source('analysis/make.distance.matrix.R')
source('analysis/fit.plot.mds.R')
```

## Jaccard index as a metric of similarity/dissimilarity

- <https://en.wikipedia.org/wiki/Jaccard_index>
- $J(A,B)=\frac{A \cap B}{A \cup B}$
- $0 <= J(A,B) <= 1$

```{r}
jaccard.all.df <- read.csv(file='analysis/data/jaccard.csv', header=TRUE)
next.group.i <- 1
```

## Hierarchical Cluster of Jaccard indices on `r levels(jaccard.all.df$Group)[next.group.i]`
<div class="centered">
```{r, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
d <- make.distance.matrix(levels(jaccard.all.df$Group)[next.group.i])
fit.plot.mds(group=group, d)
next.group.i <- next.group.i + 1
```
</div>

## Hierarchical Cluster of Jaccard indices on `r levels(jaccard.all.df$Group)[next.group.i]`
<div class="centered">
```{r, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
d <- make.distance.matrix(levels(jaccard.all.df$Group)[next.group.i])
fit.plot.mds(group=group, d)
next.group.i <- next.group.i + 1
```
</div>

## Hierarchical Cluster of Jaccard indices on `r levels(jaccard.all.df$Group)[next.group.i]`
<div class="centered">
```{r, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
d <- make.distance.matrix(levels(jaccard.all.df$Group)[next.group.i])
fit.plot.mds(group=group, d)
next.group.i <- next.group.i + 1
```
</div>

## Hierarchical Cluster of Jaccard indices on `r levels(jaccard.all.df$Group)[next.group.i]`
<div class="centered">
```{r, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
d <- make.distance.matrix(levels(jaccard.all.df$Group)[next.group.i])
fit.plot.mds(group=group, d)
next.group.i <- next.group.i + 1
```
</div>

## Hierarchical Cluster of Jaccard indices on `r levels(jaccard.all.df$Group)[next.group.i]`
<div class="centered">
```{r, echo=FALSE, echo=FALSE, warning=FALSE, message=FALSE}
d <- make.distance.matrix(levels(jaccard.all.df$Group)[next.group.i])
fit.plot.mds(group=group, d)
next.group.i <- next.group.i + 1
```
</div>
