---
title: "Analyzing differences in Jaccard indices"
author: "Rick Gilmore"
date: "`r Sys.time()`"
output: 
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: no
---


# Purpose

This document summarizes Rick Gilmore's analysis of participant sorting data.

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "img/",
  dev = "pdf",
  dpi = 300,
  fig.align = "center",
  fig.width = 7,
  fig.height = 5
)

library(tidyverse)
library(lmerTest)
```

# Analysis plan

1. Were there differences in mean *Jaccard indices* by wallpaper group?

# Import data

The Jaccard index data are found in `analysis/data/jaccard.csv`.

```{r}
jaccard_df <- readr::read_csv("analysis/data/jaccard.csv")
```

# Visualizations

```{r, jaccard-histogram}
jaccard_plot_1 <- jaccard_df %>%
  ggplot(.) +
  aes(x = Jaccard, fill = Group) +
  geom_histogram() +
  facet_grid(. ~ Group)
jaccard_plot_1
```

```{r, jaccard-boxplot}
jaccard_plot_2 <- jaccard_df %>%
  ggplot(.) +
  aes(x = Jaccard, fill = Group) +
  geom_boxplot() +
  facet_grid(Group ~ .) +
  scale_y_continuous(NULL, breaks = NULL)
jaccard_plot_2
```

# Analyses

## Using `lm`

There are no random effects in this model.

```{r}
jaccard_lm <- lm(Jaccard ~ Group, data = jaccard_df)
summary(jaccard_lm)
anova(jaccard_lm)
```

The mean Jaccard similarity indices across the exemplars was higher for P1; however, note the outliers.

